---
title: "Spatial Analysis of Houston 2021 Blackout"
description: |
  Using spatial analysis in R to explore the impacts of the February 2021 Texas storms on the greater Houston metropolitan area. Specifically looking at impacts on residential buildings and if any socioeconomic factors led to a difference in likelihood of impact. 
author:
  - name: Clarissa Boyajian
date: 2021-11-04
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r Load libraries}
library(tidyverse)
library(sf)
library(stars)
library(raster)
library(patchwork)
library(tmap)
library(rosm)
```

In February 2021, Texas was hit by three major storms, that caused a huge power crisis for residents throughout the state. In this blog, I will use spatial analysis techniques to separate out which residential buildings were impacted by the first two storms based on night light data and if any socioeconomic factors led to a difference in impact.

### Loading and Manipulating Night Light Images for Houston metropolitan area
The images are from NASA’s Level-1 and Atmosphere Archive & Distribution System Distributed Active Archive Center (LAADS DAAC). Houston sits on the border of two images, so I imported and combined the necessary images for February 7th, 2021 and February 16th, 2021. 

```{r}
read_dnb <- function(file_name) {
  # Reads the "DNB_At_Sensor_Radiance_500m" dataset from a VNP46A1 granule into a STARS object.
  # Then read the sinolsoidal tile x/y positions and adjust the STARS dimensions (extent+delta)

  # The name of the dataset holding the nightlight band in the granule
  dataset_name <- "//HDFEOS/GRIDS/VNP_Grid_DNB/Data_Fields/DNB_At_Sensor_Radiance_500m"

  # From the metadata, we pull out a string containing the horizontal and vertical tile index
  h_string <- gdal_metadata(file_name)[199]
  v_string <- gdal_metadata(file_name)[219]
  
  # We parse the h/v string to pull out the integer number of h and v
  tile_h <- as.integer(str_split(h_string, "=", simplify = TRUE)[[2]])
  tile_v <- as.integer(str_split(v_string, "=", simplify = TRUE)[[2]])

  # From the h/v tile grid position, we get the offset and the extent
  west <- (10 * tile_h) - 180
  north <- 90 - (10 * tile_v)
  east <- west + 10
  south <- north - 10

  # A tile is 10 degrees and has 2400x2400 grid cells
  delta <- 10 / 2400

  # Reading the dataset
  dnb <- read_stars(file_name, sub = dataset_name)

  # Setting the CRS and applying offsets and deltas
  st_crs(dnb) <- st_crs(4326)
  st_dimensions(dnb)$x$delta <- delta
  st_dimensions(dnb)$x$offset <- west
  st_dimensions(dnb)$y$delta <- -delta
  st_dimensions(dnb)$y$offset <- north
  
  return(dnb)
}
```

```{r, results = FALSE}
#read in each tile and give appropriate name
before_light_image1 <- "../../blog_data/spatial_analysis/VNP46A1.A2021038.h08v05.001.2021039064328.h5"
dnb_before_1 <- read_dnb(file_name = before_light_image1)

before_light_image2 <- "../../blog_data/spatial_analysis/VNP46A1.A2021038.h08v06.001.2021039064329.h5"
dnb_before_2 <- read_dnb(file_name = before_light_image2)

after_light_image1  <- "../../blog_data/spatial_analysis/VNP46A1.A2021047.h08v05.001.2021048091106.h5"
dnb_after_1 <- read_dnb(file_name = after_light_image1)

after_light_image2 <- "../../blog_data/spatial_analysis/VNP46A1.A2021047.h08v06.001.2021048091105.h5"
dnb_after_2 <- read_dnb(file_name = after_light_image2)

#Combine the tiles to create one layer for each day
combined_dnb_before <- st_mosaic(dnb_before_1, dnb_before_2)
combined_dnb_after <- st_mosaic(dnb_after_1, dnb_after_2)

#Remove older data to save memory
rm(dnb_before_1, dnb_before_2, dnb_after_1, dnb_after_2)
```

```{r fig.height=5, fig.width=3.5}
plot_combined_before <- plot(combined_dnb_before, main = "2021-02-07 Night Lights")

plot_combined_after <- plot(combined_dnb_after, main = "2021-02-16 Night Lights")
```





Disclaimer: this blog post is based on an assignment for James Frew's Spatial Analysis for Environmental Data Science course as part of the Masters of Environmental Data Science program within the Bren School of Environmental Science & Management at the University of California, Santa Barbara
