---
title: "A function I learned"
description: |
  Second Practice Blog Post
author:
  - name: Clarissa Boyajian
    url: {}
date: 08-10-2021
output:
  distill::distill_article:
    self_contained: false
preview: images/palmerpenguin.png
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(palmerpenguins)
```

## Using `ggplot` and `geom_point()` to Explore Palmer Penguin Data

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species)) +
  scale_color_hue(l = 50, c = 65) +
  labs(title = "Palmer Penguins", 
       x = "Flipper Length (mm)", 
       y = "Body Mass (g)",
       color = "Species")
```

Here I used the `ggplot` `geom_point()` function to create a scatter plot that compares the flipper lengths (mm) to the body mass (g). I added `scale_color_hue(l = 50, c = 65)` to make the species indication colors darker. I also added a title and updated labels for the the axis. I was unable to figure out how to change the legend title.

As you can see, the flipper length (mm) and body mass (g) are positively correlated in all penguin species.

